# Data Visualization

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, fig.height = 4, fig.width = 6)
```

These notes correspond to material from Chapter 3 of R for [Data Science](https://r4ds.had.co.nz) by Wickham and Grolemund.


## Loading and Exploring Data

#### Loading Data and R packages  {-}

We'll load an R package called `tidyverse`. This is a poweful R package that is useful for creating graphics, and wrangling data, and many other tasks.  

If the package is not already installed, you will need to install it, using the instructions in Assignment 0. Once installed, we load the package using the `library()` command. You will need to do this everytime you start a new R session and want to use the package. 

```{r}
library(tidyverse)
```

We will work with a dataset containing information on movies released in 2018. Data were collected from https://www.imdb.com/. 

We first read in the data from my github repository, where it is posted, and assign it the name "movies"

```{r}
movies <- read_csv("https://raw.githubusercontent.com/AndrewjSage/andrewjsage.github.io/master/Datasets/movies2018.csv")
```

#### Exploring the Data {-}

The `head()` command displays the first 6 rows of the dataset

```{r}
head(movies)
```

The `glimpse()` command returns a summary of the information contained in the dataset, including variable types. 

```{r}
glimpse(movies)
```

* Categorical variables take on groups or categories, rather than numeric values. In R, these might be coded as logical `<logi>`, character `<chr>`, factor `<fct>` and ordered factor `<ord>`.  


* Quantitative variables take on meaningful numeric values. These include numeric `<num>`, integer `<int>`, and double `<dbl>`.   

The `summary()` command returns a summary of each variable in the dataset

```{r}
summary(movies)
```

###  Creating Graphics 

The `ggplot` command is used to create graphics in `tidyverse`. 

### Example 1.1.1 Basic Scatterplot {-}

We create a scatterplot of IMDB score vs. runtime for 2018 movies. 

```{r}
ggplot(data = movies) + 
  geom_point(mapping = aes(x = run_time, y = IMDB_score))
```

### Example 1.1.2 Scatterplot with text labels {-}

Out of curiousity, let's add the names of the films that scores especially high or low. 

We'll see the `filter` command more later on. 

```{r, fig.height=6}
ggplot(data = movies) + 
  geom_point(mapping = aes(x = run_time, y = IMDB_score)) +
  geom_text(data=movies %>% filter(IMDB_score > 8 | IMDB_score < 4 ), 
            aes(x = run_time, y = IMDB_score, label = title, vjust=-1), 
            size = 2, color="red")
```

### Example 1.1.3 Scatterplot with color, shape, and size {-}


We can use color, shape, and size to convey information about other variables. 

We'll use color for genre, shape for MPAA rating, and size for gross revenue generated.

Note that shape should be used for factor variables, and size for numeric variables. Color may be used for either. 

```{r}
ggplot(data = movies) + 
  geom_point(mapping = aes(x = run_time, y = IMDB_score, 
                           color=genre, shape=MPAA_rating, size=gross))
```


### Example 1.1.4 Changing Axis Labels and Hiding Legend {-}

```{r}
ggplot(data = movies) + 
  geom_point(mapping = aes(x = run_time, y = IMDB_score, 
                           color=genre, shape=MPAA_rating, size=gross)) +
  xlab("Runtime (in minutes)")+ ylab("IMDB Score (out of 10)")
```

### Practice 1.1.5  {-}

Fill in the blanks to create a scatterplot that shows IMDB score on the y-axis and critics score (`meta_score`) on the x-axis. Use color to denote MPAA_rating, and size to represent the number of votes on IMDB (`vote`). Add appropriate axis lables. 

When you are done, remove `eval=FALSE` from the braces in the chunk heading, so your code will run. 

```{r, eval=FALSE}
ggplot(data = ___________) + 
  geom_point(mapping = aes(x = ____________, y = ___________, 
                           color=__________, size=________)) +
  xlab("_____________")+ylab("__________")
```




## More Graphics 

### Visualizing Quantitative Data (Continued)

### Example 1.2.1: Faceting {-}

`facet_wrap()` allows us to display a graph in multiple panels. You should only facet on categorical variables. 

```{r}
ggplot(data = movies) + 
  geom_point(mapping = aes(x = run_time, y = IMDB_score, color=MPAA_rating, size=gross)) +
  facet_wrap(~ genre, nrow = 2) +
  xlab("Runtime (in minutes)")+
  ylab("IMDB Score (out of 10)")
```

`facet_grid()` allows us to display a graph using a grid of categorical variables.  

```{r, fig.width=10}
ggplot(data = movies) + 
  geom_point(mapping = aes(x = run_time, y = IMDB_score, size=gross)) +
  facet_grid( MPAA_rating ~ genre) +
  xlab("Runtime (in minutes)")+
  ylab("IMDB Score (out of 10)")
```


### Example 1.2.2: Histograms {-}

Histograms are useful for visualizing a quantitative variable.

```{r}
ggplot(data = movies) + 
  geom_histogram(mapping = aes(x = run_time), color="white", fill="blue" ) + 
  xlab("Run time (in minutes")
```


### Example 1.2.3: Boxplots {-}

Boxplots can be used to graph a single quantitative variable alone, or plotted against a categorical variable.

`coordflip()` flips the x and y axes. 

```{r}
ggplot(data = movies, mapping = aes(x = MPAA_rating, y = run_time)) + 
  geom_boxplot() + xlab("MPAA Rating") + ylab("Run time (in minutes)") + 
  coord_flip()
```

### Practice 1.2.4 {-}

Fill in the blanks to create a histogram to visualize IMDB scores. 

```{r, eval=FALSE}
ggplot(data = _________) + 
  geom_histogram(mapping = aes(x = _________), color="white", fill="blue" ) + 
  xlab("_____________")
```

### Practice 1.2.5 {-}

Use faceting to display the histogram of scores for each genre separately. 

```{r, eval=FALSE}
ggplot(data = _________) + 
  geom_histogram(mapping = aes(x = _________), color="white", fill="blue" ) +
  xlab("____________") + 
  facet_wrap(~_______, nrow=2) 
```

### Practice 1.2.6 {-}

Fill in the blanks to create a boxplot to visualize IMDB scores by genre. 

```{r, eval=FALSE}
ggplot(data = movies, mapping = aes(x = ___________, y = ___________)) + 
  geom_boxplot() + xlab("____________") + ylab("_______________") +
  coord_flip()
```

### Visualizing Categorical Data

### Example 1.2.7: Bar Graph {-}

Bar graphs are used to display frequencies for categorical variables. 

```{r}
ggplot(data = movies) + 
  stat_count(mapping = aes(x = genre), fill="blue") + 
  xlab("MPAA Rating")
```

### Example 1.2.8 Reordering Bars  {-}

Notice that, by default, R orders the bars alphabetically. We might wish to specify a different order, perhaps using decreasing bar height. 

The `fct_infreq` command automatically orders factor variables according to frequency. See [this link](https://forcats.tidyverse.org/reference/fct_inorder.html) for more ordering options.  

```{r}
ggplot(data = movies) + 
  stat_count(mapping = aes(x = fct_infreq(genre)), fill="blue") + 
  xlab("MPAA Rating")
```


### Example 1.2.9 Relative Frequency Bar Graph {-}

```{r}
ggplot(data = movies) + 
  stat_count(mapping = aes(x = genre, y = ..prop.., group = 1), fill="blue") + 
  xlab("MPAA Rating")
```

### Practice 1.2.10 {-}

Fill in the blanks to create a bar graph displaying the number of films with each MPAA rating

```{r, eval=FALSE}
ggplot(data = ___________) + 
  stat_count(mapping = aes(x = ____________), fill="blue") + 
  xlab("_____________")
```


### Practice 1.2.11 {-}

Copy/paste your code from above and modify it to display the bars in descending order of heights. 

```{r}
# type code here
```

### Practice 1.2.12 {-}

Copy/paste your code from above and modify it to display proportions, rather than counts on the y-axis. 

```{r}
# type code here
```


### Example 1.2.13 Stacked Bar Graph {-}

A stacked bar graph allows us to display information about two or more categorical variables. 

```{r}
ggplot(data = movies) + 
  stat_count(mapping = aes(x = genre, fill=MPAA_rating) ) + 
  xlab("Genre")
```

### Example 1.2.14 Stacked Bar Graph with Equal Heights {-}

When comparing percentages, it is useful to have bars of the same height. We can do this using `position="fill"`. 

```{r, fig.width=10}
ggplot(data = movies) + 
  stat_count(mapping = aes(x = genre, fill=MPAA_rating), position="fill" ) + 
  xlab("Genre")
```


### Example 1.2.15 Side-by-Side Bar Graph {-}

Sometimes it is helpful to have side-by-side bars of different heights. This is done using `position="dodge"`.

```{r, fig.width=10}
ggplot(data = movies) + 
  stat_count(mapping = aes(x = genre, fill=MPAA_rating), position="dodge" ) + 
  xlab("Genre")
```

### Example 1.2.16 Height Specified by Variable {-}

Let's narrow down our dataset to the 5 most popular films, using the `filter()` command. 

```{r, fig.width=10}
movies_subset <- movies %>% filter(title %in% c("Black Panther", "Avengers: Infinity War", "Incredibles 2", "Jurassic World: Fallen Kingdom", "Aquaman"))
```

Let's create a bar graph that displays the number of IMDB users who rated each of these films (i.e. that `vote`) variable. Now, the height of the bars is given by variable in the dataset, rather than the number of occurances of a factor variable. We use the `stat=identity` command to tell R that the bar height is determined by the y-variable, rather than freqency. 

```{r, fig.width=10}
ggplot(data=movies_subset) +
  geom_bar(mapping=aes(x=title, y=vote), stat = "identity", fill="blue") + 
  xlab("Title") + ylab("Votes")
```

We can use `fct_reorder()` to reorder the bars by decreasing height. See [link](https://forcats.tidyverse.org/reference/fct_reorder.html) for more. 

```{r, fig.width=10}
ggplot(data=movies_subset) +
  geom_bar(mapping=aes(x=fct_reorder(.f=title, .x=vote, .desc=TRUE), y=vote), stat = "identity", fill="blue") +
  xlab("Title") + ylab("Votes")
```

### Practice 1.2.17 Height Specified by Variable {-}

Copy/paste the code from Example 1.2.16 and modify it to display the 5 movies' gross revenue (`gross`) on the y-axis. Arrange in descending order. 

```{r}
# type code here
```



## Advanced Data Graphics

Let's take a look at some graphics that go beyond the basics. We'll recreate two graphics that orignally appeared in a fivethirtyeight.com blogpost [How to tell Someone's Age when All You Know is Her Name](https://fivethirtyeight.com/features/how-to-tell-someones-age-when-all-you-know-is-her-name/)  

This example comes from Chapter 3 of [Modern Data Science with R](https://beanumber.github.io/mdsr2e/), by Baumer, Kaplan, and Horton. 

```{r}
library(tidyverse)
library(babynames)
library(mdsr)
```

### Babynames dataset {-}

```{r}
Babynames_data <- make_babynames_dist()
head(Babynames_data)
```

### Example 1.3.1: Age Distribution for a Given Name  {-}

### Choosing Name and Gender {-}

```{r}
Babyname <- Babynames_data %>% filter(name == "Joseph" & sex == "M")

name_plot <- ggplot(data=Babyname, aes(x=year)) + 
  geom_bar(aes(y=est_alive_today/1000), fill="#b2d7e9", color="white", stat="identity")
name_plot
```

See the [R Cookbook](http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/) for more information on color palette. 

### Setting Axis Limits {-}

```{r}
name_plot <- name_plot + ylim(c(0,42))
name_plot
```

### Adding Line  {-}

```{r}
name_plot <- name_plot + geom_line(aes(y=count_thousands), size=2)
name_plot
```

### Axis Labels {-}

```{r}
name_plot <- name_plot + 
  xlab("Year") + ylab("Number of people (in thousands)")
name_plot
```

### Adding Text {-}

```{r}
name_plot <- name_plot + 
  ggtitle("Age Distribution of American Boys Named Joseph") + 
  geom_text(x=1935, y=40, label="Number of Josephs\nborn each year") +
  geom_text(x=1915, y=13, label="Number of Josephs \nborn each year\nestimated to be alive on \n 1/1/2014", color="#b2d7e9") 
name_plot
```


### Adding Median Age  {-}

```{r}
#install.packages("Hmisc")

library(Hmisc)
#wtd.quantile <- Hmisc::wtd.quantile
median_yob <- with(Babyname, wtd.quantile(year, est_alive_today, probs=0.5))
median_yob
```

```{r}
name_plot <- name_plot + 
  geom_bar(stat="identity", color="white", fill="#008fd5", aes(y = ifelse(year==median_yob, est_alive_today / 1000 , 0)))
name_plot
```

```{r}
name_plot <- name_plot + 
  geom_text(x=2003, y=40, label="The median \nliving Joseph \nis 37 years old", color="darkgray") + 
  geom_curve(x=1995, xend=1974, y=40, yend=24, arrow=arrow(length=unit(0.3, "cm")), curvature=0.5) 
name_plot
```

### Example 1.3.2: Common Female Names  {-}

### Some data prep  {-}

```{r}
com_fem <- Babynames_data %>%
  # hack added in 2nd edition
  filter(n > 100) %>%
  filter(sex == "F") %>% 
  group_by(name) %>%
  mutate(wgt = est_alive_today / sum(est_alive_today)) %>%
  summarise(
    N = n(), est_num_alive = sum(est_alive_today),
    q1_age = wtd.quantile(age_today, est_alive_today, probs = 0.25, na.rm = TRUE),
    median_age = wtd.quantile(age_today, est_alive_today, probs = 0.5, na.rm = TRUE),
    q3_age = wtd.quantile(age_today, est_alive_today, probs = 0.75, na.rm = TRUE)
  ) %>%
  arrange(desc(est_num_alive)) %>%
  head(25)
```

### Setup the Plot  {-}

We put the names along the x-axis in decreasing order of median age.

```{r}
w_plot <- ggplot(data = com_fem, 
                 aes(x = reorder(name, -median_age), 
  y = median_age)) + 
  xlab(NULL) + ylab("Age (in years)") + 
  ggtitle("Median ages for females with the 25 most common names")
w_plot
```

### Add Linerange and color  {-}

```{r}
w_plot <- w_plot + 
  geom_linerange(aes(ymin = q1_age, ymax = q3_age), 
  color = "#f3d478", size = 10, alpha = 0.8)
w_plot
```

### Add points at median  {-}

```{r}
w_plot <- w_plot +
  geom_point(fill = "#ed3324", colour = "white", size = 4, shape = 21)
w_plot
```

### Add Labels and Flip Coordinates  {-}

```{r}  
w_plot <- w_plot + 
  geom_point(aes(y = 55, x = 24), fill = "#ed3324", colour = "white", 
    size = 4, shape = 21) + 
  geom_text(aes(y = 58, x = 24, label = "median")) + 
  geom_text(aes(y = 26, x = 16, label = "25th")) + 
  geom_text(aes(y = 51, x = 16, label = "75th percentile")) + 
  geom_point(aes(y = 24, x = 16), shape = 17) + 
  geom_point(aes(y = 56, x = 16), shape = 17) +
  coord_flip()  
w_plot
```


