# Programming


```{r}
library(tidyverse)
library(rvest)
```


This module pertains to chapters 17 -21 in Wickham and Grolemund's text. We'll learn about basic programming in R. 

```{r}
library(tidyverse)
```

## Vectors

Let's investigate how R performs operations on vectors. We'll start by generating four vectors. 

```{r}
x <- 1:10    # integers from 1 through 10
y <- seq(from = 2, to =29, by =3) # count from 2 to 29 by increments of 3
z <- rep(7, 5) # repeat 7, 5 times
w <- sample(1:10, 6) #randomly select six integers from 1 to 10
```

```{r}
x
```


```{r}
y
```


```{r}
z
```


```{r}
w
```

### Practice 5.1.1 {-}

Perform each of the following operations in R. Write a sentence explaining what each operation does. If anything surprises you, note that. 

1)
```{r}
typeof(x)
```

2)
```{r}
x + 5
```

3)
```{r}
x + y
```

4)
```{r}
x * y
```

5)
```{r}
x + z
```

6)
```{r}
x + w
```


7)
```{r}
sort(x)
```

8)
```{r}
length(x)
```

9)
```{r}
mean(x)
```

10)
```{r}
x > 5
```

11)
```{r}
mean(x > 5)
```

12)
```{r}
x %% 5 == 0 
```

13)
```{r}
y[1]
```

14)
```{r}
y[3]
```

15)
```{r}
y[(c(3,1))]
```

16)
```{r}
y[1:5]
```

17)
```{r}
y[-2]
```

18)
```{r}
y[x %% 5 == 0]
```


## Functions

A function takes in one or more inputs and returns one or more outputs. 

### Example 5.2.1: Function with Real Number Inputs  {-}

Write a function called `sum_of_squares` that takes in two real numbers x and y, and returns the sum of the squares of these numbers. 

```{r}
sum_of_squares <- function(x, y){
  # inputs are real numbers x and y
  # output is sum of squares of x and y
  z <- x^2 + y^2
  return(z)
}
```

```{r}
sum_of_squares(x=3, y=7)
```

```{r}
sum_of_squares(x= c(1, 3), y=c(2, 4))
```

### Example 5.2.2: Function with Vectors and Dataframes as Inputs  {-}

The `starwars` dataset contains information on characters that appear in the Star Wars movies. 

```{r}
data(starwars)
head(starwars)
```


```{r}
ggplot(data=starwars, aes(x=mass)) + geom_histogram() + ggtitle("Mass Distribution")
```

```{r}
starwars %>% arrange(desc(mass))
```


```{r}
ggplot(data=starwars, aes(x=height)) + geom_histogram() + ggtitle("Height Distribution")
```

```{r}
starwars %>% arrange(desc(height))
```

```{r}
ggplot(data=starwars, aes(x=birth_year)) + geom_histogram() + ggtitle("Birth Year Distribution")
```

```{r}
starwars %>% arrange(desc(birth_year))
```


```{r}
mean(starwars$mass, na.rm=TRUE)
```

Why might this not be a good representation of the average mass of star wars characters?

A **trimmed mean** is a version of the mean that excludes the largest and smallest m numbers (where m is an integer). Calculate trimmed means for mass, height, and birth year that exclude the maximum and minumum values. 

Why is the following code sub-optimal
```{r}
#mass calculation
(sum(starwars$mass, na.rm=TRUE) - min(starwars$mass, na.rm=TRUE) - max(starwars$mass, na.rm=TRUE)) / (sum(!is.na(starwars$mass)) - 2)
```

```{r}
#height calculation
(sum(starwars$height, na.rm=TRUE) - min(starwars$height, na.rm=TRUE) - max(starwars$height, na.rm=TRUE)) / (sum(!is.na(starwars$height)) - 2)
```

```{r}
#age calculation
(sum(starwars$birth_year, na.rm=TRUE) - min(starwars$birth_year, na.rm=TRUE) - max(starwars$birth_year, na.rm=TRUE)) / (sum(!is.na(starwars$birth_year)) - 2)
```


Write a function that calculates the trimmed mean, excluding the maximum and minimum value for a given vector. Exclude NA's from the calculation.

```{r}

trimmed_mean <- function(x){
  # input a vector x
  # output the mean, excluding largest and smallest values, and NA's
  
  tm <- (sum(x, na.rm=TRUE) - max(x, na.rm=TRUE) - min(x, na.rm=TRUE)) / (sum(!is.na(x))-2)
  return(tm)
}
```

```{r}
trimmed_mean(x=starwars$mass)
trimmed_mean(x=starwars$height)
trimmed_mean(x=starwars$birth_year)
trimmed_mean(y)
trimmed_mean(c(3,5))
```


### Example 5.2.3  {-}

Write a function that inputs a dataframe and returns every fifth row. 

```{r}
return_every_fifth <- function(df){
  # take in a dataframe
  # return rows 5, 10, 15, etc.
  
  keep <- seq( from = 5, to = nrow(df), by = 5 )
  df1 <- df[ keep , ]
  return(df1)
}

```


```{r}
return_every_fifth(df=starwars)
```

### Practice 5.2.4   {-}

Write functions to perform each of the following. Exclude NA's from the calculation. Test the functions on the starwars dataset, where appropriate. 

a) Input two real numbers x and y, and return the sum of the absolute values. 

```{r}
# type here
```


b) Input two real numbers x and y and return the larger of the two. 

```{r}
# type here
```


c) Input a vector and calculate the range of values (max - min). 

```{r}
# type here
```


d) Input a vector and calculate the variance. Note: $\frac{1}{n-1}\sum(x-\bar{x})^2$. Do not use the built-in function `var`.    

```{r}
# type here
```


e) Input a dataframe and return 5 rows, selected at random. Hint: use `sample`. 

```{r}
# type here
```


### Example 5.2.5    {-}

Write a function that takes in a function defined on the real numbers, and returns a graph, using a viewing window specified by the user. 

```{r}
plot_function <- function(f, xrange, yrange){
  # input f is a function on reals
  # input xrange is a vector for x-axis
  # input yrange is a vector for y-axis
  p <- ggplot(data = data.frame(xrange, yrange), aes(x=xrange, y=yrange)) + stat_function(fun=f) + xlim(xrange) + ylim(yrange)
return(p)
}
```

```{r}
square <- function(x){
  #input a real number
  #return its square
  return(x^2)
}
```


```{r}
plot <- plot_function(f=square, xrange = c(-2, 2), yrange = c(0,5) ) 
```

```{r}
plot
```


### Example 5.2.6: Approximating Area   {-}

The standard normal probability density function is given by $f(x)=\frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}.$ We define the function in R. 

```{r}
normal_density <- function(x){
#input real number x
#return standard normal density at x   
  return(1/sqrt(2*pi)*exp(-x^2/2))
  }
```

```{r}
plot_function(f=normal_density, xrange = c(-3, 3), yrange = c(0,0.5) ) 
```

Note: this function is actually already available via the built-in function `dnorm`

```{r}
plot_function(f=dnorm, xrange = c(-3, 3), yrange = c(0,0.5) ) 
```


This function does not have a closed-form anti-derivative, so we cannot find areas under the curve using standard calculus techniques. Instead, we can use simulation to approximate area. This technique is called **Monte Carlo Integration**. 

We'll use the `runif` function, which generates a specified number of random numbers (n), between a minimum (min) and maximum (max) value. 

```{r}
x <- runif(n=1, min=0, max=1)
y <- runif(n=1, min=0, max=1)
x
y
```



#### 1. Describe the steps you could use to approximate the area under the standard normal curve between $x=-1$ and $x=1$, using simulation. 


#### 2. Write a function that performs these steps. Allow the user to input the following arguments. Can you make the function return both an approximation of the area, and also a visual representation?

Arguments:     
n: number of points to simulate
f: function to integrate
x1: minimum of x range    
x2: maximim of x range      
y1: minimum of y range    
y2: maximum of y range    


```{r}
simulate_area <- function(n, f, x1, x2, y1, y2){
  # f is function to find area under
  # xmin and xmax give range of x-values
  # ymin and ymax give range of y-values
x <- runif(n, x1, x2) 
y <- runif(n, y1, y2)
under <- y < f(x)
df <- data.frame(x, y, under)
p <- ggplot(data = df, mapping = aes(x = x, y=y)) + geom_point(aes(color = under)) + stat_function(fun=f)
area_est <- mean(under) * (x2 - x1) * (y2 - y1)
return(list(p, area_est))
}
```

```{r}
simulate_area(n=100, f=normal_density, x1=-1, x2=1, y1=0, y2=0.5)
```

```{r}
simulate_area(n=1000, f=normal_density, x1=-1, x2=1, y1=0, y2=0.5)
```

```{r}
simulate_area(n=1000, f=normal_density, x1=-1, x2=1, y1=0, y2=1)
```


### 3. Test the function using n=10, n=1000, n=10000, and n=100000.   {-}

## Iteration    

### Example 5.3.1: For Loop for Monte Carlo Integration  {-}

We suspected that using larger values of $n$ would give us more precise approximations. Write a function that runs 1,000 simulations, using samples of size specified by the user, and returns the resulting estimates. Test the function using n=100, 1000, and 10000, and create histograms displaying the distribution of the resulting estimates. What can we conclude?

```{r}
RepeatSim <- function(sampsize){
Results <- vector("double", 1000)
for( i in 1:1000){
  Results[i] <- simulate_area(n=sampsize, f=dnorm, x1=-1, x2=1, y1=0, y2=1)[[2]]
}
return(Results)
}
```


a) n=100
```{r}
Results_n100 <- RepeatSim(sampsize=100)
```

```{r}
ggplot(data=data.frame(Results_n100), aes(x=Results_n100)) + geom_histogram() 
```

b) n=1,000

```{r}
Results_n1000 <- RepeatSim(sampsize=1000)
```

```{r}
ggplot(data=data.frame(Results_n1000), aes(x=Results_n1000)) + geom_histogram() 
```


c) n=10,000

```{r}
Results_n10000 <- RepeatSim(sampsize=10000)
```


```{r}
ggplot(data=data.frame(Results_n10000), aes(x=Results_n10000)) + geom_histogram() 
```



### Example 5.3.2: Iteration on Dataframes    {-}

```{r}
data(diamonds)
diamonds <- diamonds %>% select_if(is.numeric) # create a copy of the dataset
diamonds <- data.frame(diamonds)
```

In many situations, we will want to perform the same function to multiple columns of a dataframe. For example, it is often helpful to standardize variables before entering them into a statistical model or a machine learning algorithm. 

Note: standardizing means subtracting the mean, and dividing by the standard deviation. Thus, a standardized score (or z-score) tells us how many standard deviations an observation lies above or below the mean. 

We could easily write our own function to standardize, but the `scale` function in R does this already. 


Create a new dataframe containing standardized versions of all numeric variables in the diamonds dataset, using each of the following three techniques: 


1. Using `transmute`
```{r}
diamonds_scaled <- diamonds %>% transmute(
  carat = scale(carat),
  depth = scale(depth), 
  table = scale(table), 
  price = scale(price), 
  x = scale(x), 
  y = scale(y),
  z = scale(z)
  )

head(diamonds_scaled)
```


2. Using a for loop
```{r}
diamonds_scaled2 <- diamonds
for(i in seq_along(diamonds)){
  diamonds_scaled2[,i] <- scale(diamonds[,i])
}

head(diamonds_scaled2)
```

Note: the textbook recommends using `seq_along(diamonds)` in place of `1:ncol(diamonds)`. These functions are equivalent, except `seq_along` better handles zero-length vectors.  

3. Using the `apply()` function

The `apply()` function applies a function over all rows or columns of a dataframe. The `MARGIN` argument specifies whether the function should be applied to the rows(1) or columns(2) of the dataframe. 

```{r}
diamonds_scaled3 <- apply(diamonds, MARGIN=2, FUN=scale)

head(diamonds_scaled3)
```

Note `apply` is one of a family of functions including `lapply`, `vapply`, and `mapply` which work on different kinds of structures, such as lists and vectors. The `purrr` package, which is part of tidyverse includes a set of mapping functions, discussed in section 21.5 of the text, which perform similar tasks. 

### Practice 5.3.3  {-}

We previously wrote a function called `trimmed_mean()` to calculate the mean of a vector, excluding the smallest and largest value. 

```{r}
trimmed_mean <- function(x){
  # input a vector x
  # output the mean, excluding largest and smallest values, and NA's
  tm <- (sum(x, na.rm=TRUE) - max(x, na.rm=TRUE) - min(x, na.rm=TRUE)) / (sum(!is.na(x))-2)
  return(tm)
}
```

Calculate the trimmed mean of each column of the diamonds dataset using 

a) a for loop

```{r}
# type here
```

b) the `apply` function

```{r}
#type here
```


### Example 5.3.4: Loop for scraping   {-}

We used the following code to scrape the 50 highest grossing 2019 movies from imdb.com.

```{r}
library(rvest)
```


```{r}
page <- read_html("https://www.imdb.com/search/title/?title_type=feature&release_date=2019-01-01,2019-12-31&sort=boxoffice_gross_us,desc&start=101&ref_=adv_nxt")

title <- page %>%
  html_nodes(".lister-item-header a") %>%
  html_text()

rating <- page %>%
  html_nodes(".ratings-bar") %>%
  html_text() %>%
  str_remove_all(" ") %>%
  str_remove_all("\n") %>%
  str_remove_all("\"") %>%
  str_split("Ratethis") %>%
  map(1, .default=NA) %>%
  as.numeric()

runtime <- page %>%
  html_nodes(".lister-item-header+ .text-muted") %>%
  html_text() %>%
  as.character() %>% 
  str_replace("\n            \n", "\n           NA \n |" )  %>%
  str_split("\\|") %>%
  map(2, .default=NA) %>% 
  str_remove_all("min") %>%
  trimws() %>%
  as.numeric()
  

genre <- page %>%
  html_nodes(".lister-item-header+ .text-muted") %>%
  html_text() %>%
  as.character() %>% 
  str_replace("\n            \n", "\n           NA \n |" )  %>%
  str_split("\\|") %>%
  map(4, .default=NA) %>%
  trimws()

votes <- page %>%
  html_nodes(".sort-num_votes-visible span:nth-child(2)") %>%
  html_text() %>%
  str_remove_all(",") %>%              # remove all commas in each entry
  as.numeric() %>% as.numeric()

gross <- page %>%
  html_nodes(".sort-num_votes-visible") %>%
  html_text() %>%
  str_replace("\n            \n", "\n           NA \n |" ) %>%
  str_split("\\|") %>%
  map(2, .default=NA) %>% 
  str_remove_all("Gross:\n") %>%
  str_remove_all("\\$") %>%
  str_remove_all("M") %>%
  trimws() %>% 
  as.numeric()

MPAA_rating <- page %>%
  html_nodes(".lister-item-header+ .text-muted") %>%
  html_text %>%
  str_replace("\n            \n", "\n           NA \n |" ) %>%
  str_split("\\|") %>%
  map(1, .default=NA) %>% 
  trimws() %>%
  as.character()

metascore <- page %>%
  html_nodes(".ratings-bar") %>%
  html_text %>%
  str_remove_all(" ") %>%
  str_remove_all("\n") %>%
  str_split("X") %>%
  map(2, .default=NA) %>%
  str_remove_all("Metascore") %>%
  str_trim() %>%
  as.numeric()

movies <- tibble(title, genre, rating, runtime, metascore, MPAA_rating, gross, votes)
```

We can incorporate this code into a function and a loop to scrape in multiple pages of films. 

Write a function that takes in the rank of the first movie on the page (in terms of gross), and returns a dataframe with the 50 movies on that page. 

Use `sprintf` to read in a url with a variable url. Place `%d` in a url in place of a variable numeric (double) variable. Similarly, use `%s` in place of a character variable. 

```{r}
Read_page <- function(pagestart){
  
url <- sprintf("https://www.imdb.com/search/title/?title_type=feature&release_date=2019-01-01,2019-12-31&sort=boxoffice_gross_us,desc&start=%d&ref_=adv_nxt", pagestart)

page <- read_html(url)

title <- page %>%
  html_nodes(".lister-item-header a") %>%
  html_text()

rating <- page %>%
  html_nodes(".ratings-bar") %>%
  html_text() %>%
  str_remove_all(" ") %>%
  str_remove_all("\n") %>%
  str_remove_all("\"") %>%
  str_split("Ratethis") %>%
  map(1, .default=NA) %>%
  as.numeric()

runtime <- page %>%
  html_nodes(".lister-item-header+ .text-muted") %>%
  html_text() %>%
  as.character() %>% 
  str_replace("\n            \n", "\n           NA \n |" )  %>%
  str_split("\\|") %>%
  map(2, .default=NA) %>% 
  str_remove_all("min") %>%
  trimws() %>%
  as.numeric()
  

genre <- page %>%
  html_nodes(".lister-item-header+ .text-muted") %>%
  html_text() %>%
  as.character() %>% 
  str_replace("\n            \n", "\n           NA \n |" )  %>%
  str_split("\\|") %>%
  map(4, .default=NA) %>%
  trimws()

votes <- page %>%
  html_nodes(".sort-num_votes-visible span:nth-child(2)") %>%
  html_text() %>%
  str_remove_all(",") %>%              # remove all commas in each entry
  as.numeric() %>% as.numeric()

gross <- page %>%
  html_nodes(".ghost~ .text-muted+ span") %>%
  html_text() %>%
  str_remove("\\$") %>%  #note \\ is required before special characters
  str_remove("M") %>%
  as.numeric()

MPAA_rating <- page %>%
  html_nodes(".lister-item-header+ .text-muted") %>%
  html_text %>%
  str_replace("\n            \n", "\n           NA \n |" ) %>%
  str_split("\\|") %>%
  map(1, .default=NA) %>% 
  trimws() %>%
  as.character()

metascore <- page %>%
  html_nodes(".ratings-bar") %>%
  html_text %>%
  str_remove_all(" ") %>%
  str_remove_all("\n") %>%
  str_split("X") %>%
  map(2, .default=NA) %>%
  str_remove_all("Metascore") %>%
  str_trim() %>%
  as.numeric()

movies <- tibble(title, genre, rating, runtime, metascore, MPAA_rating, gross, votes)
  
return(movies)
}

```

Write a loop that collects data on the 150 highest grossing movies, and collects them in a dataframe. 

```{r}
movies <- NULL

for( i in c(1, 51, 101)){
  movies1 <- Read_page(pagestart = i)
  movies <-rbind(movies, movies1)
}

```

```{r}
head(movies)
```

```{r}
tail(movies)
```


```{r}
dim(movies)
```


